Dynamic Load Balancing for\\Heterogeneous Systems

### 2011-12-05 paper/paper.tex
Added to GPU and CPU architectures.

\section{Architecture Overview}
In this section we discuss the different architectures of the experimental
machines under test.  We will discuss CPU, GPU, and APU architectures in
terms of major differences that affect the performance of the device.

@@

The use of the Single Instruction Multiple Data (SIMD) paradigm enables increased
performance on these devices by increasing data throughput for data parallel
applications.  However, optimal performance of these types of instructions often
involves complex compilation techniques and hand-tuning of the assembly.

@@

Groups of processing units, Streaming Multiprocessors (SMs), share a single instruction
dispatch unit and operate in lock step.  The threads themselves run on each SM
in a group, called a warp, typically of 32 threads.  Because of these architectural
features divergent branching in a warp (caused by if-else clauses) greatly degrade application
performance.

>>>>

\section{Architecture Overview}
In this section we discuss the different architectures of the experimental
machines under test.  We will discuss CPU, GPU, and APU architectures in
terms of major differences that affect the performance of the device.  We also
briefly discuss the impacts of different architectures on the implementations
of a CPU+GPU co-processing system.

@@

In addition, the CPU typically has a deep pipeline to increase throughput of the system.
Longer and longer pipelines increase the utilization of every element found on the CPU,
further increasing latency-critical performance.  Performance of this pipeline is in danger
of being flushed on braching codes, but advanced branch prediction and speculative execution
are able to greatly reduce the amount of times where a pipeline flush must occur.  Out-of-order
execution of instructions in the program also improves performance for single threads on the
CPU.

The use of the Single Instruction Multiple Data (SIMD) paradigm enables increased
performance on these devices by increasing data throughput for data parallel
applications.  However, optimal performance of these types of instructions often
involves complex compilation techniques and hand-tuning of the assembly.  Most applications
are not able to see as substantial a speedup from this process as they would if the application
were really able to process 4 threads at once.

@@

Each processor on the GPU is very dumb and primitive compared to the processors on the CPU.
There is no speculative execution, deep pipelines, out-of-order exeuction, or branch prediction
on this device.  Instead, the space saved by not having these features is instead used to
cram more and more compute cores onto the device.  This strategy of having a large number of
very simple cores goes back to the orginial architectual decisions for the GPU.  Graphics applications
are relatively straight-forward and require a lot of math and very little complex coding.
Therefore, for graphics, there is little performance lost for not having these features, whereas
for GPGPU applications these features can greatly diminish overall kernel performance.

Groups of processing units, Streaming Multiprocessors (SMs), share a single instruction
dispatch unit and operate in lock step.  The threads themselves run on each SM
in a group, called a warp, typically of 32 threads.  Because of these architectural
features divergent branching in a warp (caused by if-else clauses) greatly degrade application
performance.  Unlike the CPU, the GPU is unable to avoid warp-level divergence of
the different threads, and therefore must always take the penalty of braching.
